Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate

### 2015-07-10 masters_paper/paper.tex
Adding the paper

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=1in]{geometry}
\linespread{1.5}
\pdfpagewidth 8.5in
\pdfpageheight 11in

@@

\subsection{Knockoff Filter}
    Describe the original knockoff filter,

@@

    Simulations with deterministic knockoffs reveal that they don't perform well in L1 regularized logistic regression. Even when $X_i$ is a null predictor of $y$, the $X_i$ still tend to enter the model prior to $\tilde X_i$. The issue is that even when $x\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, the $\tilde X_i$ do not fit a normal distribution. This can be seen below in figure 1, where normal Q-Q plots of simulated variables $X_i$ and corresponding knockoffs $\tilde X_i$ are provided.
    \begin{figure}[h!]
        \includegraphics[width=20cm]{images/normalQQ}
        \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
    \end{figure}

    This can be seen from producing qq plots of $X_i$ vs $\tilde X_i$ for each $i$. Of course, when $X$ is a binary vector, $\tilde X$ completely doesn't match its distribution, causing the original $X$ to beat the knockoffs into the model. This indicates that a new method of generating $\tilde X$ must be created to control FDR via knockoffs with regularized logistic regression.

\section{Random Bernoulli Knockoffs}
    My idea is to generate $\bf \tilde x$ randomly such that, approximately, $\bf \tilde x\sim \bf x$. In particular, both variables should have similar marginal densities, expectations, and second moments. However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\mathbf{\tilde x}^T\mathbf x\st \mathbf x) = \Sigma-\mathbf s$, where $\|\diag(\Sigma)-s\|$ is small. In the general case, this is likely infeasible; however, if $\bf x$ is a binary vector, as is often the case, we know we are dealing with a much more limited class of random variables, and it should be possible to randomly generate $\bf \tilde x\st \bf x$ so as to have the desired properties. At worst, this method will provide a suitable replacement for deterministic $\bf \tilde x$ for use with LASSO, and if we are lucky, it will work reasonably for other regularized GLMs.

\section{Random Bernoulli Generation}
    Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values. A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$. Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag(M)$. For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm^T\succeq0$, and
    \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
    for all $i\neq j$\footnote{``On parametric families for sampling binary data with specified mean and correlation" - http://arxiv.org/abs/1111.0576}. Given a qualifying $M$, or observed $X$, there are a few ways of generating more random $\bf x$.

\subsection{Gaussian Copula Family}
    Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $\mathbf z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties. There are a number of ways to do this\footnote{``On the Generation of Correlated Artificial Binary Data" - http://epub.wu.ac.at/286/1/document.pdf}\footnote{``On parametric families for sampling binary data with specified mean and correlation"}, but it turns out that there is only certain to exist a working $\Sigma$ in the bivariate case.

\subsection{$\mu$-Conditionals family}
    There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as

@@

\section{Generating Knockoffs}
    My method for generating binary knockoffs broadly involves two steps:
    \begin{enumerate}
        \item I use either the equal correlation method or the SDP method described in the original knockoff paper to find $\mathbf s$ such that $\|\diag(\hat\Sigma)-\mathbf s\|$ is small and 
        \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag(\mathbf s )\\ \hat\Sigma - \diag(\mathbf s) & \hat\Sigma \end{array}\right] \succeq 0 \]
        I use a subscript $L$ for large to denote items associated with the joint distribution of $[\mathbf x \; \mathbf{\tilde x}]$. If $\mathbf m_L = \E\left([\mathbf x \; \mathbf{\tilde x}]\right)^T = [\mathbf m \; \mathbf m]^T$, then the desired cross moment matrix of the joint distribution is 
        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
        To ensure that this is a valid cross moment matrix for a binary random vector is that 
        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq M_{L,ij} \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
        I've built a check into the code for this, but in practice it shouldn't be a worry. This condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$. Since the we are either keeping the value $M_{L,ij}$ from a valid cross moment matrix or minimizing $\vert M_{L,ij} - \mathbf m_{L,j}\mathbf m_{L,i}\vert$, it would be very surprising if this condition was violated.

@@

>>>>

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\linespread{1.5}
\pdfpagewidth 8.5in
\pdfpageheight 11in

@@

\subsection{Knockoff Filter}
    Describe the original knockoff filter, and introduce some notation. Original=deterministic

@@

    Simulations with deterministic knockoffs reveal that they don't perform well in L1 regularized logistic regression. Even when $X_i$ is a null predictor of $y$, the $X_i$ still tend to enter the model prior to $\tilde X_i$. The issue is that even when $x\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, the $\tilde X_i$ do not fit a normal distribution. This can be seen below in figure 1, where normal Q-Q plots of simulated variables $X_i$ and corresponding knockoffs $\tilde X_i$ are provided. Of course, when $X$ is a binary vector, $\tilde X$ completely doesn't match its distribution.
    \begin{figure}[h!]
        \includegraphics[width=20cm]{images/normalQQ}
        \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
    \end{figure}
    These

\section{Random Bernoulli Knockoffs}
    My solution to the issues with the

     generate $\bf \tilde x$ randomly such that it has the desired knockoff properties in expectation. In particular, both  should have similar marginal densities, expectations, and second moments. However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\mathbf{\tilde x}^T\mathbf x\st \mathbf x) = \Sigma-\mathbf s$, where $\|\diag(\Sigma)-s\|$ is small. In the general case, this is likely infeasible; however, if $\bf x$ is a binary vector, as is often the case, we know we are dealing with a much more limited class of random variables, and it should be possible to randomly generate $\bf \tilde x\st \bf x$ so as to have the desired properties. At worst, this method will provide a suitable replacement for deterministic $\bf \tilde x$ for use with LASSO, and if we are lucky, it will work reasonably for other regularized GLMs.

\subsection{Random Bernoulli Generation}
    Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values. A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$. Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag(M)$. For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm^T\succeq0$, and
    \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
    for all $i\neq j$\footnote{``On parametric families for sampling binary data with specified mean and correlation" - http://arxiv.org/abs/1111.0576}. Given a qualifying $M$, or observed $X$, there are a few ways of generating more random $\bf x$.

\subsubsection{Gaussian Copula Family}
    Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $\mathbf z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties. There are a number of ways to do this\footnote{``On the Generation of Correlated Artificial Binary Data" - http://epub.wu.ac.at/286/1/document.pdf}\footnote{``On parametric families for sampling binary data with specified mean and correlation"}, but it turns out that there is only certain to exist a working $\Sigma$ in the bivariate case.

\subsubsection{$\mu$-Conditionals family}
    There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as

@@

\subsection{Generating Binary Knockoffs}
    My method for generating binary knockoffs broadly involves two steps:
    \begin{enumerate}
        \item I use either the equal correlation method or the SDP method described in the original knockoff paper to find $\mathbf s$ such that $\|\diag(\hat\Sigma)-\mathbf s\|$ is small and 
        \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag(\mathbf s )\\ \hat\Sigma - \diag(\mathbf s) & \hat\Sigma \end{array}\right] \succeq 0 \]
        I use a subscript $L$ for large to denote items associated with the joint distribution of $[\mathbf x \; \mathbf{\tilde x}]$. If $\mathbf m_L = \E\left([\mathbf x \; \mathbf{\tilde x}]\right)^T = [\mathbf m \; \mathbf m]^T$, then the desired cross moment matrix of the joint distribution is 
        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
        To ensure that this is a valid cross moment matrix for a binary random vector is that 
        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq M_{L,ij} \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
        I've built a check into the code for this, but in practice it shouldn't be a worry. This condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$. Since the we are either keeping the value $M_{L,ij}$ from a valid cross moment matrix or minimizing $\vert M_{L,ij} - \mathbf m_{L,j}\mathbf m_{L,i}\vert$, it would be very surprising if this condition was violated.

@@

\section{Bernoulli Knockoff Performance}
For each of thses, exaimine these situatations:
\begin{itemize}
    \item $X$ arises out of the assumed Isling model, with no high order interactions. 
    \item $X$ does not arise out of this model, and does have higher order interactions. Could possibly model this by drawing the vector $\mathbf x$ or subsets of it as multinomial with probabilities from a Dirichlet distribution.
    \item $X$ is from a real world data set, likely something in genetics.
    \item $y$ is normally distributed around $X\beta$.
    \item $y$ is drawn from some other distribution, possibly skewed, heavy tailed, or light tailed.
    \item $y$ is from a real world data set.
\end{itemize}
\subsection{Gram Matrix Comparison}
\subsection{Binary vs. Determinist Knockoffs in LASSO}
\subsection{Binary Knockoffs in other Regularlized GLMS}

\section{Discussion}
\subsection{Areas for Further Work}
